### 2017-02-24 2017mergingmerges/section/XX-Related.tex
Notes on synchronous/passive dataflow and scenario-aware dataflow

\subsection{Scratch: dataflow}
Synchronous languages (LUSTRE, Lucy-n, SIGNAL) view streams as -- abstractly -- functions from time to values.
(Or time to maybe values, if they allow holes).
This forces them into an inherently \emph{push} based system, but what is pushing here is the procession of time.
If you have two streams, you cannot pull from one and the other: when time advances, it advances for all streams.
Thus, synchronous languages cannot encode value-dependent pulling - nor can they encode append.
Append could be seen as a special kind of value-dependent access pattern, if you treat the end of the stream as a special kind of value.

@@

Do not to confuse regular dataflow, also known as \emph{synchronous data flow}, with synchronous languages.
Regular dataflow graphs are dataflow graphs where each edge is annotated with the rate.

Dataflow languages:
\begin{itemize}
\item Lucid Synchrone
\item LUSTRE \cite{halbwachs1991synchronous}.
Dataflow language: causal analysis, clock calculus, synchronous.
Uses more or less syntactic equality to check equivalence of clocks.
\item Lucy-n: extension of LUSTRE with periodic clocks to ensure static scheduling \cite{mandel2010lucy}, \cite{benveniste2003synchronous}
\item SIGNAL \cite{le2003polychrony}
causal, clocks, synchronous
\item StreaMIT \cite{thies2002streamit}:
Regular/synchronous dataflow, ie statically known rates.
Only allows limited splits and joins: round robin and duplication for splits, round robin and combination for joins. 
Does not support fully general graphs - instead using combinators to introduce a (split/join) and a combinator for a feedback loop.
\item Prelude: extension of LUSTRE, seems very similar to Lucy-n \cite{pagetti2011multi}
\item Ptolemy (II)
\end{itemize}

Models:
\begin{itemize}
\item Synchronous dataflow (SDF):
all rates are statically known,
\item Cyclo-static dataflow (CSDF) where the number of produced/consumed tokens varies periodically, sounds very similar to Lucy-n's periodic clocks \cite{mandel2010lucy}
\item Boolean dataflow (BDF), generally undecidable to statically schedule, so must fall back to dynamic scheduling \cite{buck1993scheduling}
\item Integer dataflow (IDF), (also falls back to dynamic) \cite{buck1994static}
Clustering to find control structure: tries to recover structures like ifs, cases and loops from the dataflow graph.
Extends previous BDF \cite{buck1993scheduling} with loops that repeat some number of times, and multi-way ifs.
Very rigid and only supports limited control flow structures.
Unclear how a merge or append would be clustered by this system.
\item Parameterized dataflow (PDF),  \cite{bhattacharya2001parameterized}
\item Schedulable parametric dataflow (SPDF),  \cite{fradet2012spdf}
\end{itemize}

>>>>

\subsection{Scratch: dataflow}
Synchronous languages (LUSTRE, Lucy-n, SIGNAL) view streams as -- abstractly -- functions from time to values.
(Or time to maybe values, if they allow holes).
Synchrony is described as ``time advances in lockstep with one or more clocks'' \cite{benveniste2003synchronous}.
This forces them into an inherently \emph{push} based system, but what is pushing here is the procession of time.
If you have two streams, you cannot pull from one and the other: when time advances, it advances for all streams.
Thus, synchronous languages cannot encode value-dependent pulling - nor can they encode append.
Append could be seen as a special kind of value-dependent access pattern, if you treat the end of the stream as a special kind of value.

Contrast between our case, where we have \emph{active} processes which perform the pulling and pushing, with synchronous cases which are more passive, in that they don't request input - the values trickle in from outside.

@@

Do not to confuse regular dataflow, also known as \emph{synchronous data flow}, with synchronous languages.
Regular dataflow graphs are dataflow graphs where each edge is annotated with the rate.
The closest thing to fusion in dataflow computation is static scheduling, where the order in which different actors are fired at runtime is chosen statically.
This is somewhat similar to fusion, in fusion we are also interested in merging the separate actors into a single one that computes them all.
Choosing a static schedule is required to fuse actors together, but it alone is not fusion.
In our system, we are performing both static scheduling and fusion at the same time.

Dataflow languages:
\begin{itemize}
\item Lucid Synchrone
``takes advantage of the (first-order) functional aspect of Lustre so as to generalize it to higher order constructs in an ML-like style. In particular, the Lustre clock calculus is extended and inferred as an ML-type system.'' \cite{benveniste2003synchronous}
\item LUSTRE \cite{halbwachs1991synchronous}.
Dataflow language: causal analysis, clock calculus, synchronous.
Uses more or less syntactic equality to check equivalence of clocks.
\item Lucy-n: extension of LUSTRE with periodic clocks to ensure static scheduling \cite{mandel2010lucy}
\item SIGNAL \cite{le2003polychrony}
causal, clocks, synchronous
\item StreaMIT \cite{thies2002streamit}:
Regular/synchronous dataflow, ie statically known rates.
Only allows limited splits and joins: round robin and duplication for splits, round robin and combination for joins. 
Does not support fully general graphs - instead using combinators to introduce a (split/join) and a combinator for a feedback loop.
\item Ptolemy (II)
\end{itemize}

Models:
\begin{itemize}
\item Synchronous dataflow (SDF):
all rates are statically known,
\item Cyclo-static dataflow (CSDF) where the number of produced/consumed tokens varies periodically, sounds very similar to Lucy-n's periodic clocks \cite{mandel2010lucy}
\item Boolean dataflow (BDF), generally undecidable to statically schedule, so must fall back to dynamic scheduling \cite{buck1993scheduling}
\item Integer dataflow (IDF), (also falls back to dynamic) \cite{buck1994static}
Clustering to find control structure: tries to recover structures like ifs, cases and loops from the dataflow graph.
Extends previous BDF \cite{buck1993scheduling} with loops that repeat some number of times, and multi-way ifs.
Very rigid and only supports limited control flow structures.
Unclear how a merge or append would be clustered by this system.
\item Parameterized dataflow (PDF),  \cite{bhattacharya2001parameterized}
\item Schedulable parametric dataflow (SPDF),  \cite{fradet2012spdf}
\item Scenario aware dataflow (FSM-SADF) \cite{stuijk2011scenario}
Somehow use a separate finite state machine, which can control the rates of dataflows.
Apparently more expressive/succinct than boolean dataflow (BDF) but also easier to analyze.
`General SADF' must fall back to runtime scheduling, while FSM-SADF can be done statically.
In FSM-SADF, however, the currently executing scenario can only be switched in between iterations of the whole graph.
So after the whole graph executes one scenario, the FSM is used to find the next scenario to execute.
This means that the currently executing scenario cannot depend on values read during the current scenario - which rules out paritioning, for example, because the choice of whether to emit on the true or false branch cannot be made until after the value has been emitted.
It is possible to indirectly encode this by storing the value to emit in a buffer, and deferring the emit until next scenario - however this loses locality benefits of fusion.
General SDF could describe a partition, but uses runtime scheduling.
The focus is more about analysing worst-case execution time in order to ensure real time deadlines can be met \cite{van2015scenario}.

Basically each node of the FSM is a dataflow graph. After the current dataflow graph is executed, the FSM can take a step based on the output of that graph. Then the graph at the new node is executed, and so on.
Because the FSM is separate to the dataflow graph, it is not as composable as a combinator-based approach: adding the equivalent to a new combinator could require modifying many of the dataflow graphs, and the FSM, in non-trivial ways.
\end{itemize}
